---
title: "How To Use CGMissingDataR"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{How To Use CGMissingDataR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# CGMissingDataR

CGMissingDataR supports two related continuous glucose monitoring missing-data
workflows:

* benchmarking imputation methods by masking known glucose values, and
* imputing glucose values that are already missing in user data.

The current workflow functions accept a raw timestamp column and create
`TimeSeries` and `TimeDifferenceMinutes` internally. You can choose which model
outputs to return with `models`. The default is `models = "mice_only"` for a
quick MICE-only run. Use `models = "all"` to run MICE-only, Random Forest, kNN,
XGBoost, LightGBM, and ARIMA, or pass a subset such as
`models = c("mice_only", "rf")`.

## Installation

Before installation, make sure the modeling dependencies are available:

```r
install.packages(c(
  "FNN", "ranger", "mice", "xgboost", "lightgbm", "forecast",
  "CGManalyzer", "lifecycle"
))
```

Install the development version of CGMissingDataR from GitHub:

``` r
devtools::install_github("ZhangLabUKY/CGMissingDataR")
```

Load the package:

```{r setup}
library(CGMissingDataR)
```

## Example Data

The package includes two small multi-subject CGM example datasets. Both contain
raw timestamps, not precomputed `TimeSeries` or `TimeDifferenceMinutes` columns.
Those generated time features are created inside the imputation functions.

`CGMExampleData` is intended for benchmarking. The glucose values are complete,
so the benchmark function can mask known values and compare imputed values
against the truth.

```{r example-data-complete}
data("CGMExampleData")

data.frame(
  Dataset = "CGMExampleData",
  Rows = nrow(CGMExampleData),
  Subjects = length(unique(CGMExampleData$USUBJID)),
  MissingGlucose = sum(is.na(CGMExampleData$LBORRES)),
  MissingPercent = round(mean(is.na(CGMExampleData$LBORRES)) * 100, 1)
)

head(CGMExampleData)
```

`CGMExampleData2` has the same columns but includes deterministic missing
glucose values. It is intended for demonstrating imputation of actual missing
target values.

```{r example-data-missing}
data("CGMExampleData2")

data.frame(
  Dataset = "CGMExampleData2",
  Rows = nrow(CGMExampleData2),
  Subjects = length(unique(CGMExampleData2$USUBJID)),
  MissingGlucose = sum(is.na(CGMExampleData2$LBORRES)),
  MissingPercent = round(mean(is.na(CGMExampleData2$LBORRES)) * 100, 1)
)

head(CGMExampleData2)
```

## Benchmark Known Glucose Values

Use `run_comprehensive_imputation_benchmark()` when glucose values are known
and you want to mask a portion of them to compare model performance.

This example masks 5% of the glucose rows using random masking and runs all
available methods. Smaller model iteration counts are used so the example stays
reasonably quick.

```{r benchmark-example}
benchmark_out <- run_comprehensive_imputation_benchmark(
  CGMExampleData,
  target_col = "LBORRES",
  feature_cols = c("AGE", "hba1c"),
  id_col = "USUBJID",
  time_col = "Time",
  time_format = "yyyy:mm:dd:hh:nn",
  mask_rates = 0.05,
  mask_type = "random",
  models = "all",
  rf_n_estimators = 25,
  xgb_nrounds = 25,
  lgb_nrounds = 25
)
```

The `results` table contains one row per selected model and mask rate.

```{r benchmark-results}
benchmark_out$results
```

Each selected model also gets its own completed data frame in
`benchmark_out$imputed_data`.

```{r benchmark-imputed-names}
names(benchmark_out$imputed_data)
```

The model-specific data frames include the original observed glucose value,
whether the row was masked, and the model-specific completed value.

```{r benchmark-rf-preview}
head(
  benchmark_out$imputed_data$rf[
    ,
    c(
      "USUBJID", "Time", ".Masked", "ObservedValue", "ImputedValue",
      "TimeSeries", "TimeDifferenceMinutes"
    )
  ]
)
```

## Impute Real Missing Glucose Values

Use `run_missing_glucose_imputation()` when your data already contains missing
glucose values. This function returns imputed values but does not calculate
accuracy metrics, because the true values for those missing rows are unknown.

This example uses `CGMExampleData2`, where 50 of the 500 glucose rows are
missing.

```{r real-imputation-example}
impute_out <- run_missing_glucose_imputation(
  CGMExampleData2,
  target_col = "LBORRES",
  feature_cols = c("AGE", "hba1c"),
  id_col = "USUBJID",
  time_col = "Time",
  time_format = "yyyy:mm:dd:hh:nn",
  models = "all",
  rf_n_estimators = 25,
  xgb_nrounds = 25,
  lgb_nrounds = 25
)
```

The `summary` table reports the missingness seen after preprocessing.

```{r real-imputation-summary}
impute_out$summary
```

The returned `imputed_data` object is a named list with one completed data frame
per selected method.

```{r real-imputation-names}
names(impute_out$imputed_data)
```

In real-imputation output, the original target column is left unchanged. That
means `LBORRES` is still `NA` where glucose was originally missing, while
`ImputedValue` contains the completed glucose value.

```{r real-imputation-rf-preview}
rf_missing_rows <- impute_out$imputed_data$rf[
  impute_out$imputed_data$rf$.Missing,
  c(
    "USUBJID", "Time", ".Missing", "LBORRES", "ObservedValue",
    "ImputedValue", "TimeSeries", "TimeDifferenceMinutes"
  )
]

head(rf_missing_rows)
```
