---
title: "How To Use CGMissingDataR"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{How To Use CGMissingDataR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# CGMissingDataR

CGMissingDataR imputes missing glucose values in continuous glucose monitoring
(CGM) data. The main public workflow is `run_missing_glucose_imputation()`,
which accepts a raw timestamp column, creates `TimeSeries` and
`TimeDifferenceMinutes` internally, and returns completed values from the
selected imputation models.

The timestamp column can be stored in common formats such as
`2020:01:16:00:00`, `2020-01-16 00:00:00`, `2020/01/16 00:00:00`, or as a
`POSIXct` column. The function standardizes those timestamps before calling
`CGManalyzer::timeSeqConversion.fn()`.

Use `models = "mice_only"` for a quick MICE-only run. Use `models = "all"` to
run MICE-only, Random Forest, kNN, XGBoost, LightGBM, and ARIMA, or pass a
subset such as `models = c("mice_only", "rf")`.

Simulation-based benchmarking is kept as an internal development and validation
workflow.

## Installation

Before installation, make sure the modeling dependencies are available:

```r
install.packages(c(
  "FNN", "ranger", "mice", "xgboost", "lightgbm", "forecast",
  "CGManalyzer", "lifecycle"
))
```

Install the development version of CGMissingDataR from GitHub:

``` r
devtools::install_github("ZhangLabUKY/CGMissingDataR")
```

Load the package:

```{r setup}
library(CGMissingDataR)
```

## Example Data

`CGMExampleData` is a small multi-subject CGM dataset with raw timestamps and
deterministic missing glucose values. It does not include `TimeSeries` or
`TimeDifferenceMinutes`; those generated time features are created inside the
imputation function.

```{r example-data}
data("CGMExampleData")

data.frame(
  Dataset = "CGMExampleData",
  Rows = nrow(CGMExampleData),
  Subjects = length(unique(CGMExampleData$USUBJID)),
  MissingGlucose = sum(is.na(CGMExampleData$LBORRES)),
  MissingPercent = round(mean(is.na(CGMExampleData$LBORRES)) * 100, 1)
)

head(CGMExampleData)
```

## Impute Missing Glucose Values

Use `run_missing_glucose_imputation()` when your data already contains missing
glucose values. This function returns imputed values but does not calculate
accuracy metrics, because the true values for those missing rows are unknown.

This example uses all available methods. Smaller model iteration counts are
used so the example stays reasonably quick.

```{r real-imputation-example}
impute_out <- run_missing_glucose_imputation(
  CGMExampleData,
  target_col = "LBORRES",
  feature_cols = c("AGE", "hba1c"),
  id_col = "USUBJID",
  time_col = "Time",
  models = "all",
  rf_n_estimators = 25,
  xgb_nrounds = 25,
  lgb_nrounds = 25
)
```

The `summary` table reports the missingness seen after preprocessing.

```{r real-imputation-summary}
impute_out$summary
```

The returned `imputed_data` object is a named list with one completed data frame
per selected method.

```{r real-imputation-names}
names(impute_out$imputed_data)
```

In the returned data, the original target column is left unchanged. That means
`LBORRES` is still `NA` where glucose was originally missing, while
`ImputedValue` contains the completed glucose value.

```{r real-imputation-rf-preview}
rf_missing_rows <- impute_out$imputed_data$rf[
  impute_out$imputed_data$rf$.Missing,
  c(
    "USUBJID", "Time", ".Missing", "LBORRES", "ObservedValue",
    "ImputedValue", "TimeSeries", "TimeDifferenceMinutes"
  )
]

head(rf_missing_rows)
```
